\ZpL: a p-adic precision package

### 2018-01-29 ZpL/ZpL.tex
Done part of the examples

\begin{itemize}
\item determinant
\item characteristic polynomial
\end{itemize}

@@

\begin{itemize}
\item gcd of polynomials
\end{itemize}

@@

\subsubsection{$p$-adic differential equations}

>>>>

In linear algebra, determinant and 
characteristic polynomial are notoriously
hard to compute: see \cite{caruso-roe-vaccon:15, caruso-roe-vaccon:17}. We illustrate this with the following
example (using a division-free algorithm
for the computation of the characteristic polynomial) :

\bigskip

\begin{tabular}{rl}
\cIn
 & \verb?Q2 = Zp(2, prec=10, print_mode="digits")? \\
 & \verb?MS = MatrixSpace(Q2,3,3)? \\
 & \verb?Mat = MS.random_element()? \\
 & \verb?chi = Mat.charpoly()? \\
 & \verb?deto = Mat.determinant()? \\
 & \verb?deto? \\
\cZpCR
 & \verb?...000100000100? \\
\cZpFP
 & \verb? ...11010101010? \\
\cZpLC
 & \verb?  ...0110110001? \\
\cZpLF
 & \verb?...001110111100? \\
\end{tabular}

\begin{tabular}{rl}
\cIn
 & \verb?chi? \\
\cZpCR
 & \verb?(...0000000001)*x^3 + (...1001111111)*x^2 ?\\
 & \verb?+ (...01110101110)*x + (...111011111100)? \\
\cZpFP
 & \verb?...0000000001*x^3 + ...1101001001*x^2 + ?\\
 & \verb?...0100110011*x + ...00101010110? \\
\cZpLC
 & \verb?...0000000001*x^3 + ...1011000111*x^2 + ?\\
 & \verb?...0011101000*x + ...1001001111? \\
\cZpLF
 & \verb?...0000000001*x^3 + ...1111000101*x^2 + ?\\
 & \verb?...0011111111*x + ...110001000100? \\
\end{tabular}

@@

Our package can be applied to
computation over $p$-adic polynomials.
A natural example is that of the computation of GCD,
whose stability has been studied in \cite{caruso-2017}.

A naive implementation of the Euclide algorithm
can produce different behaviour depending 
on the type of implementation of the field
of $p$-adic coefficients.

\begin{tabular}{rl}
\cIn
 & \verb?Z2 = Zp(2, prec=15, print_mode="digits")? \\
  & \verb?S.<x> = PolynomialRing(Z2)? \\
 & \verb?P = random_element(S, degree=10, prec=5)? \\
 & \verb?Q = random_element(S, degree=10, prec=5)? \\
 & \verb?D = x^5 + random_element(S, degree=4, prec=8)? \\
 & \verb?def euclidean(A,B):? \\
 & \verb?    while B != 0:? \\
 & \verb?        A, B = B, A % B? \\
 & \verb?    return A.monic()? \\
\end{tabular}

\bigskip

\begin{tabular}{rl}
\cIn
  & \verb?euclidean(D*P, D*Q)? \\
\cZpCR
 & \verb?(...1)*x^9 + (0)*x^8 + (...,.1)*x^7 ?\\
 & \verb?+ (...,.1)*x^6 + (...1)*x^5 + (...1)*x^4?\\
 & \verb? + (...,.1)*x^3 + (0)*x^2 + (...,.1)*x + (...10)? \\
\cZpFP
 & \verb?...000000000000000000000000000001? \\
\cZpLC
 & \verb?...1*x^5+...100*x^3+...100*x^2+...1*x+...1? \\
\end{tabular}

With high probability, $P$ and $Q$ are coprime, implying that the gcd of $DP$ is $DQ$ is $D$.  
However, we observe that we do not always get this expected result.
More precisely, ZpCR often founds a gcd of higher degree.
Indeed the Euclidean algorithm stops prematurely because the software believes that some early remainder vanishes due to the lack of precision.  
The situation with ZpFP is the exact opposite:
since the precision is too high (always the maximal cap), 
the software sometimes does not notice that the remainder vanishes.
Hence, it happens that the Euclidean algorithm 
obtains a last non-zero remainder of degree less than that of
$D.$

@@

\subsection{Gröbner bases}
Our package can be applied 
on complex computations like that of
Gröbner bases using generic
Gröbner bases algorithms.

\begin{tabular}{rl}
\cIn
 & \verb?Q2 = Qp(2, prec=50, print_mode="digits")? \\
 & \verb?R.<x,y,z> = PolynomialRing(Q2, order = 'invlex')? \\
 & \verb?F = [Q2(2,10)*x + Q2(1,10)*z, ? \\
 & \verb?     Q2(1,10)*x^2 + Q2(1,10)*y^2 - Q2(2,10)*z^2,? \\
 & \verb?     Q2(4,10)*y^2 + Q2(1,10)*y*z + Q2(8,10)*z^2]? \\
 & \verb?from sage.rings.polynomial.toy_buchberger?\\
 &\verb?  import buchberger_improved? \\
 & \verb?g = buchberger_improved(ideal(F)); g.sort()? \\
 & \verb?g? \\
\cZpCR
 & \verb?[x^3, x*y + ...1100010*x^2, y^2 + ...11001*x^2,? \\
 & \verb? z + ...0000000010*x]? \\
\cZpLC
 & \verb?[[x^3, x*y + ...111100010*x^2, y^2 + ...1111111001*x^2,? \\
 & \verb? z + ...0000000010*x]]? \\
\end{tabular}

As we can see, some loss in precision occurs in the
Buchberger algorithm.
Had we used \cZpFP , many padded zeroes or one would have appeared.

\subsection{$p$-adic differential equations}
